Aggressive Scheduling for Numerical Programs

### 2012-04-30 report/draft.tex
Add new reference. Add abstract.

\title{Aggressive Scheduling for Numerical Programs}
\author{Richard Veras, Flavio Cruz, Berkin Akin}
\date{\today}
\documentclass[12pt]{article}

@@

@@

However, there are certain classes of algorithms were this approach falls short in improving
the performance of the target code. One good example is numerical code. Expert programmers
are able to take a very na\"\i ve implementation and then fine tune it to take advantage of the
micro-architecture features of the target machine. Such features include things like software
pipelining (XXX include citation here) (for improving the performance of Out-Of-Order execution),
unrolling (by knowing how big are the memory caches of the processor) and pre-fetching.
Unfortunately, compilers are unable to perform such optimizations due
to some loss of generality and also because most optimizations performed by a compiler
cannot see the forest for the trees, and thus can't take into account a complete numerical algorithm.
This problem is further emphasized by the fact that expert programmers test and tune various permutations
of their code and its scheduling of instructions, a task that compilers are usually unable to do because
they would have to guess the structure of the input data and then generate and run small instances of the problem
to find an optimal solution.

@@

@@

+( ) We want to show the breakdown of the cycles for each test and where the
     stalls were occurring.
+( ) What is our testbed? What machines do we use?

\subsection{Simple Scalar}
\begin{verbatim}
+ What is simplescalar and why are we using it?

+ We want to be able to determine the types of systems that will benefit the
  most from this kind of scheduling.

+ What parameters will we tune?

\section{Experimental Evaluation}
\begin{verbatim}
+ What algorithms are we using?
+ What systems are we using?

>>>>

\title{Aggressive Scheduling for Numerical Programs}
\author{Richard Veras, Flavio Cruz, Berkin Akin}
\date{\today}
\documentclass[10pt]{article}

@@

\begin{abstract}
   Certain classes of algorithms are still hard to optimize by modern compilers
   due to the differences between architecture features.
   Numerical algorithms, for example, perform better when taking advantage of features
   such as software pipelining and loop unrolling. Expert programmers are usually able to
   get the last drop of performance from the hardware, however, compilers cannot do the same
   for loss of generality. The SPIRAL system bridges this gap by generating machine specific code
   from high level algorithmic descriptions. We used the SPIRAL system to implement software
   pipelining for numerical algorithms with XXX (talk about scheduling). Our results show that
   ....
\end{abstract}

@@

However, there are certain classes of algorithms where this approach falls short in improving
the performance of the target code. One good example is numerical code. Expert programmers
are able to take a very na\"\i ve implementation and then fine tune it to take advantage of the
micro-architecture features of the target machine. Such features include things like software
pipelining~\cite{Rau:81} (for improving the performance of Out-Of-Order execution by interleaving loop iterations),
unrolling (by knowing how big are the memory caches of the processor) and pre-fetching.
Unfortunately, compilers are unable to perform such optimizations due
to some loss of generality and also because most optimizations performed by a compiler
cannot see the forest for the trees, and thus can't take into account a complete numerical algorithm.
This problem is further emphasized by the fact that expert programmers test and tune various permutations
of their code and its scheduling of instructions, a task that compilers are usually unable to do because
they would have to guess the structure of the input data and then generate and run small instances of the problem
to find an optimal solution.

@@

\lstset{basicstyle=\small,language=C}

@@

+( ) We want to show the breakdown of the cycles for each test and where the
     stalls were occurring.
+( ) What is our testbed? What machines do we use?
\end{verbatim}

\subsection{Simple Scalar}

\begin{verbatim}
+ What is simplescalar and why are we using it?
+ We want to be able to determine the types of systems that will benefit the
  most from this kind of scheduling.
+ What parameters will we tune?
\end{verbatim}

\section{Experimental Evaluation}

\begin{verbatim}
+ What algorithms are we using?
+ What systems are we using?
